\begin{acknowledgements}
First of all, I would like to thank my advisors: Assoc. Prof. Elif Özkırımlı for her support and patience during my studies and Assoc. Prof. Arzucan Özgür for her guidance, trust in me, and big heart that she has as an academic mom. Thank you for all the big hugs you gave virtually and personally. 

I want to express my gratitude to my jury members, Prof. Tunga Güngör and Prof. Nilgün Lütfiye Karalı, for kindly accepting to participate in the defense of this thesis and for all the comments and critics that improved my study. 

I gratefully acknowledge TÜBİTAK ARDEB 1001 Program (Project No: 119E133) and the TAM project for providing the computational resources. 

I also thank my lab-mates, whom I cannot list here since they are a lot, but I will always remember you in front of my window, full of flowers with the view of the North Campus pyramid. 

I am grateful to my family, Nuray Parlar and Mustafa Parlar, for always supporting me and giving me the opportunities that have made me who I am. Also, I would like to thank my chosen family, M.D. İremnaz Karahan and Mert Aydın for providing a mirror to my life. Last but not least,  my cat Fındık, which is like a child to me, thank you for opening a window to my heart. 

The completion of this study could not have been possible without the expertise, patience, love, support, and hugs of my dearest lifelong best friend, desk-mate, and love of my life, Rıza Özçelik. This thesis is dedicated to our new family, and I love you as much as the words in this thesis (23,235), believe it or count it.
\end{acknowledgements}